Summary of CA Rice Breeding

### 2017-02-01 output/summary.Rnw
Commit before start re-arranging from EDA to more directed

<<label=plot2, fig=TRUE, echo=FALSE>>=
vars <- c("M206","M202","M204","M205", "M209")
i <- all_vt_result$id %in% vars

boxplot(yield_lb ~ id, data = all_vt_result[i,],
        xlab = "Cultivar",
        ylab = "Yield (lbs/ac)")
@
\caption{Boxplot of yields for four common medium grain cultivars for
  the period 1995--2015.}
\end{figure}

@@

\section{Evidence of yield decreases over time}

@@

  <<fig = TRUE, echo=FALSE>>=
  plot(yield_lb ~ year, data = all_vt_result, type = "n")
  mods <- lmList(yield_lb ~ (1|year) | id, data = all_vt_result[i,])
  pred <- lapply(mods, function(x){
      data.frame(pred = unique(predict(x)),
                 year = unique(x$model$year))
  })
  sapply(seq_along(pred), function(i) points(x = pred[[i]]$year, y = pred[[i]]$pred,
                                             pch = i))
  legend("bottomright", legend = vars, pch = 1:5)

@@

<<echo=FALSE>>=
meds_split <- split(all_vt_result[i,], paste(all_vt_result[i, "year"], all_vt_result[i,"site"]))
meds_scaled <- lapply(meds_split, function(x){
    m202 <- mean(x$yield_lb[x$id == "M202"])
    x$yield_lb <- x$yield_lb - m202
    return(x)
    })
meds_scaled <- do.call(rbind, meds_scaled)
ggplot(meds_scaled, aes(y = yield_lb, x = year, group = id)) + geom_smooth(aes(color = id), method = "lm") + geom_point()

mod <- stan_glm(yield_lb ~ factor(year) + site + factor(year):site +
                    yrs_in_trial + id + id:yrs_in_trial,
                data = all_vt_result[i,], prior = normal(), cores = 2L, iter = 1000)

@ 
\begin{enumerate}
  \item Technology improvements over time, including advances in
    nutrient, pest, and water management
  \item Variability over time 
  \item Variability between sites
\end{enumerate}

@@

plot(yield_lb ~ year, all_vt_result[i,],
     xlab = "Year", ylab = "Yield (lbs/ac)")

abline(lm(yield_lb ~ year, all_vt_result[i,]))

>>>>

\begin{figure}
<<label=plot2, fig=TRUE, echo=FALSE>>=
  vars <- grepl("^M[0-9]{3}$", all_vt_result$id)
  mod_data <- all_vt_result[vars,]

  boxplot(yield_lb ~ id, data = mod_data,
          xlab = "Cultivar",
          ylab = "Yield (lbs/ac)")
@
\caption{Boxplot of yields for four common medium grain cultivars for
  the period 1995--2015.}
\end{figure}

@@

\subsection{Hypotheses}

\begin{enumerate}
  \item After a cultivar has been released, the yield decreases over
    time.
  \item Each cultivar release is recovering to the old yield, not
    increasing yield potential.
\end{enumerate}

\section{H1: Do yields decreases over time once a cultivar is released?}

\subsection{Methods}

Linear model of yield of a cultivar as a response to the number of years since it
was entered into the trials. Corrections for year, site, year:site,
and cultivar differences are made according to:

$ yield = \alpha + \beta_{years in trial j}*x_{years in trials j} +
\beta_{site} + \beta_{year} + \beta_{site:year} + \beta_{cultivar j}$

Only medium-grained, released varieties were included in the analysis,
reducing the data from \Sexpr{nrow(all_vt_result)} to
\Sexpr{nrow(mod_data)} observations. Included varieties are:
\Sexpr{paste(gsub('^M', 'M-', unique(mod_data$id)), collapse = ', ')}.

@@

% \begin{figure}

%   <<fig = TRUE, echo=FALSE>>=
%   ## plot(yield_lb ~ year, data = all_vt_result, type = "n")
%   ## mods <- lmList(yield_lb ~ (1|year) | id, data = mod_data)
%   ## pred <- lapply(mods, function(x){
%   ##     data.frame(pred = unique(predict(x)),
%   ##                year = unique(x$model$year))
%   ## })
%   ## sapply(seq_along(pred), function(i) points(x = pred[[i]]$year, y = pred[[i]]$pred,
%   ##                                            pch = i))
%   ## legend("bottomright", legend = vars, pch = 1:5)

% @

% <<echo=FALSE>>=
% ## meds_split <- split(mod_data, paste(all_vt_result[i, "year"], all_vt_result[i,"site"]))
% ## meds_scaled <- lapply(meds_split, function(x){
% ##     m202 <- mean(x$yield_lb[x$id == "M202"])
% ##     x$yield_lb <- x$yield_lb - m202
% ##     return(x)
% ##     })
% ## meds_scaled <- do.call(rbind, meds_scaled)
% ## ggplot(meds_scaled, aes(y = yield_lb, x = year, group = id)) + geom_smooth(aes(color = id), method = "lm") + geom_point()

% ## mod <- stan_glm(yield_lb ~ factor(year) + site + factor(year):site +
% ##                     yrs_in_trial + id + id:yrs_in_trial,
% ##                 data = mod_data, prior = normal(), cores = 2L, iter = 1000)

@@

\begin{enumerate}
  \item Technology improvements over time, including advances in
    nutrient, pest, and water management
  \item Variability over time 
  \item Variability between sites
\end{enumerate}

@@

plot(yield_lb ~ year, mod_data,
     xlab = "Year", ylab = "Yield (lbs/ac)")

abline(lm(yield_lb ~ year, mod_data))
